Generating Empirically Optimized Composed Matrix Kernels from MATLAB Prototypes

### 2009-02-15 doc/papers/iccs09/paper.tex
git-svn-id: https://svn.mcs.anl.gov/repos/performance/orio@446 e7ad4b5f-b827-0410-872f-f7f4bc3d1efb

\begin{table}[htb] 
%\vspace{-.1in} 
\caption{Composed BLAS operations used in our experiments.}
\label{tbl:blas-ops}
\small
\centering 
%\vspace{-.1in} 
\begin{tabular}{c c c c} 
\hline 
Name & Input & Output & Operation  \\ 
\hline 
%\hline 
%& & & \\
VADD & $\overrightarrow{w}$,$\overrightarrow{y}$,$\overrightarrow{z}$ & $\overrightarrow{x}$ &
$\overrightarrow{x} = \overrightarrow{w} + \overrightarrow{y} + \overrightarrow{z}$ \\ [1ex]
%---------------------------------------------------------------------------------
%\hline 
%& & & \\
ATAX & $A$,$\overrightarrow{x}$ & $\overrightarrow{y}$ & 
$\overrightarrow{y} = A' * (A * \overrightarrow{x})$ \\ [1ex]
%---------------------------------------------------------------------------------
%\hline 
%& & & \\
GEMVER & ${\bf A}$,$a$,$b$, &  ${\bf B}$,  & ~~${\bf B} = {\bf A} + \overrightarrow{u_1} * \overrightarrow{v_1}' + \overrightarrow{u_2} * \overrightarrow{v_2}'$\\
 & $\overrightarrow{u_1}$,$\overrightarrow{u_2}$,$\overrightarrow{v_1}$,$\overrightarrow{v_2}$, & $\overrightarrow{x}$,$\overrightarrow{w}$ & $\overrightarrow{x} = b * ({\bf B}' * \overrightarrow{y}) + \overrightarrow{z}$\\
& $\overrightarrow{y}$,$\overrightarrow{z}$ &  & $\overrightarrow{w} = a * ({\bf B} * \overrightarrow{x})$\\ [1ex]
%---------------------------------------------------------------------------------
%\hline 
BiCG Kernel & ${\bf A}$,$\overrightarrow{p}$,$\overrightarrow{r}$ & $\overrightarrow{q}$,$\overrightarrow{s}$ & $\overrightarrow{q} = {\bf A} * \overrightarrow{p}$ \\ 
& & & $\overrightarrow{s} = {\bf A}' * \overrightarrow{r}$\\

@@

The performance results (in MFLOP/s) of tuning the VADD operation on the Blue
Gene/P are given in Figure~\ref{fig:results}(a). The ``Base'' label
designates the C implementation generated by the MATLAB compiler. We also
tested the performance of an implementation that calls
\texttt{DAXPY} twice using available BLAS libraries.
Finally, we tuned the simple C loop version using Orio, with the performance
annotations previously shown in Figure~\ref{fig:orio-example}. In this
experiment, we measured the performance for both the sequential and parallel
scenarios (indicated by (S) and (P) in the legend, respectively). Even for a very simple operation such as vector addition the
compiler alone is unable to obtain the same level of performance as the
empirically tuned versions. Furthermore, as expected, the BLAS implementation
does not exploit locality and thus performed worse than the single-loop
implementation.
%of the single loop implementation due to l
% however, selecting the correct compiler
%options was crucial for achieving good performance of the base version.

@@

The Xeon performance results (in seconds) of ATAX are shown in
Figure~\ref{fig:results}(b). The Orio-tuned version that incorporates
Pluto-generated loop fusion optimizations and Orio parallelization
directives achieves the best performance for most problem sizes,
outperforming the Intel MKL version by a factor of 2 to 5.7 and the
compiler-optimized C version by a factor of 4 to 7. The optimizations
performed by both Orio versions include scalar replacement,
vectorization, and loop unroll/jam.

Figure~\ref{fig:results}(c) shows the performance (in seconds) of the GEMVER
operation on the Xeon workstation. Here we used the same Pluto and Orio
optimizations as for the ATAX example. Similarly, the parallel Orio version
achieved the best performance, although in this case the sequential Orio
version performs almost the same, suggesting that the compiler was not able
to parallelize the code very effectively. For this operation, substantial
performance differences exist between among the different BLAS versions, with
the Intel MKL version achieving performance close to that of the simple
compiler code.

The performance (in seconds) for the BiCG kernel operation is shown in
Figure~\ref{fig:results}(d). For this operation, the Pluto analysis did not
result in performance improvement. Thus we are showing the results obtained
only through Orio transformations, which included vectorization, scalar
replacement, and loop unroll/jam. Again the best performance was achieved
by the parallel Orio version, while all the BLAS versions performed worse
than the compiler-optimized C loop version.

>>>>

\begin{table}[htb] 
%\vspace{-.1in} 
\caption{Composed BLAS operations used in our experiments.}
\label{tbl:blas-ops}
\small
\centering 
%\vspace{-.1in} 
\begin{tabular}{c c c c} 
\hline 
Name & Input & Output & Operation  \\ 
\hline 
%\hline 
%& & & \\
VADD & $\overrightarrow{w}$,$\overrightarrow{y}$,$\overrightarrow{z}$ & $\overrightarrow{x}$ &
$\overrightarrow{x} = \overrightarrow{w} + \overrightarrow{y} + \overrightarrow{z}$ \\ [1ex]
%---------------------------------------------------------------------------------
%\hline 
%& & & \\
ATAX & $A$,$\overrightarrow{x}$ & $\overrightarrow{y}$ & 
$\overrightarrow{y} = A^T * (A * \overrightarrow{x})$ \\ [1ex]
%---------------------------------------------------------------------------------
%\hline 
%& & & \\
GEMVER & ${\bf A}$,$a$,$b$, &  ${\bf B}$,  & ~~${\bf B} = {\bf A} + \overrightarrow{u_1} * \overrightarrow{v_1}^T + \overrightarrow{u_2} * \overrightarrow{v_2}^T$\\
 & $\overrightarrow{u_1}$,$\overrightarrow{u_2}$,$\overrightarrow{v_1}$,$\overrightarrow{v_2}$, & $\overrightarrow{x}$,$\overrightarrow{w}$ & $\overrightarrow{x} = b * ({\bf B}^T * \overrightarrow{y}) + \overrightarrow{z}$\\
& $\overrightarrow{y}$,$\overrightarrow{z}$ &  & $\overrightarrow{w} = a * ({\bf B} * \overrightarrow{x})$\\ [1ex]
%---------------------------------------------------------------------------------
%\hline 
BiCG Kernel & ${\bf A}$,$\overrightarrow{p}$,$\overrightarrow{r}$ & $\overrightarrow{q}$,$\overrightarrow{s}$ & $\overrightarrow{q} = {\bf A} * \overrightarrow{p}$ \\ 
& & & $\overrightarrow{s} = {\bf A}^T * \overrightarrow{r}$\\

@@

The performance results of tuning the VADD operation on the Blue
Gene/P are given in Figure~\ref{fig:results}(a). The ``Base'' label
designates the C implementation generated by the MATLAB compiler. We also
tested the performance of an implementation that calls
\texttt{DAXPY} twice using available BLAS libraries.
Finally, we tuned the simple C loop version using Orio, with the performance
annotations previously shown in Figure~\ref{fig:orio-example}. In this
experiment, we measured the performance for both the sequential and parallel
scenarios (indicated by (S) and (P) in the legend, respectively). Even for a very simple operation such as vector addition the
compiler alone is unable to obtain the same level of performance as the
empirically tuned versions. Furthermore, as expected, the BLAS implementation
does not exploit locality and thus performed worse than the single-loop
implementation.
%of the single loop implementation due to l
% however, selecting the correct compiler
%options was crucial for achieving good performance of the base version.

@@

The Xeon performance results of ATAX are shown in
Figure~\ref{fig:results}(b). The Orio-tuned version that incorporates
Pluto-generated loop fusion optimizations and Orio parallelization
directives achieves the best performance for most problem sizes,
outperforming the Intel MKL version by a factor of 2 to 5.7 and the
compiler-optimized C version by a factor of 4 to 7. The optimizations
performed by both Orio versions include scalar replacement,
vectorization, and loop unroll/jam.

Figure~\ref{fig:results}(c) shows the performance of the GEMVER
operation on the Xeon workstation. Here we used the same Pluto and Orio
optimizations as for the ATAX example. Similarly, the parallel Orio version
achieved the best performance, although in this case the sequential Orio
version performs almost the same, suggesting that the compiler was not able
to parallelize the code very effectively. For this operation, substantial
performance differences exist between among the different BLAS versions, with
the Intel MKL version achieving performance close to that of the simple
compiler code.

The performance for the BiCG kernel operation is shown in
Figure~\ref{fig:results}(d). For this operation, the Pluto analysis did not
result in performance improvement. Thus we are showing the results obtained
only through Orio transformations, which included vectorization, scalar
replacement, and loop unroll/jam. Again the best performance was achieved
by the parallel Orio version, while all the BLAS versions performed worse
than the compiler-optimized C loop version.
